---
title: "ZL's Works"
format: html
editor: visual
---

## Introduction

In this project, we will explore the data of strawberry agriculture in the United States obtained from USDA. Specifically, our research will focus on the yield, the ratio of crop grown to the unit area of land, of various states from 1998 to 2024.

We will conduct some basic exploratory data analysis, trying to answer questions like:

1\) How did the yield of each state differ?

2\) How did the yield of each year differ?

3\) How did the yield of each state change across the years?

Additionally, we will see what other insights about the yield of U.S.'s strawberry agriculture we may extract from the limited data available to us.

## Set Up

```{r}
#| label: load libraries and set options
#| warning: false
#| message: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
library(ggplot2)
```

```{r}
# Define function: drop_one_value_col, which drops all columns that contain only one value
drop_one_value_col <- function(df){
  drop <- c()
  for (i in 1:ncol(df)){
    uni <- unique(df[,i])
    if (is.data.frame(uni)){
      unique_val <- nrow(unique(df[,i]))
    } else {
      unique_val <- length(unique(df[,i]))
    }

    
    
    if (unique_val == 1){
      drop <- c(drop, i)
    }
  }
  
  if (length(drop) != 0){
    df <- df[,-drop]
  }
  df
}
```

## Load and Clean the Data

```{r}
#| label: read data 

# Load the strawberrt data from csv file 
strawberry <- read.csv("strawberry_10oct25.csv") %>% 
  # Drop all single value column
  drop_one_value_col()

# Divide the data into two parts: survey & census
strawberry.sur <- strawberry %>%
  filter(Program=="SURVEY") 
strawberry.cen <- strawberry %>%
  filter(Program=="CENSUS")

```

Select data that is relevant to our research theme: yield

```{r}
# Since census data contains nothing relevant to yielding, we will focus solely on the survey data
# Select survey data 
strawberry.sur <- strawberry %>%
  filter(Program=="SURVEY") %>%
  # Select Data.Item rows that contains information relevant to yield
  filter(grepl("ACRES|PRODUCTION|YIELD", Data.Item)) %>%
  # Drop all one value columns
  drop_one_value_col()

# Clean the data 
# Remove all "STRAWBERRIES - " from the column Data.Item, since they serve no purpose 
strawberry.sur$Data.Item <- str_remove_all(strawberry.sur$Data.Item, "STRAWBERRIES - ") 
# Remove all "STRAWBERRIES, " from the column Data.Item, since they serve no purpose 
strawberry.sur$Data.Item <- str_remove_all(strawberry.sur$Data.Item, "STRAWBERRIES, ") 

# Replace "(D)" with NA
strawberry.sur$Value[strawberry.sur$Value ==" (D)"] <- NA

# Turn all values into numeric
strawberry.sur$Value <- str_remove_all(strawberry.sur$Value, ",")
strawberry.sur$Value <- as.numeric(strawberry.sur$Value)
```

Reshape the data

```{r}
strawberry.sur.wider <- strawberry.sur %>% 
  pivot_wider(
    names_from = Data.Item,
    values_from = Value
    ) %>%
  # We do not consider "FORECAST" data, filter those rows out
  filter(!grepl("FORECAST", Period)) %>%
  # Arrange rows by state
  arrange(State)  %>%
  # Drop columns that contained "TON"
  select(!contains("TON")) %>%
  # Drop all one value columns
  drop_one_value_col()
```

## Exploratory Data Analysis: Yield

Research Question: How does each state's yield differ?

```{r}
strawberry.sur.wider %>%
  ggplot(aes(x=State, y=`YIELD, MEASURED IN CWT / ACRE`, fill=State)) +
  scale_x_discrete(labels=c("CA", "FL", "NY", "NC", "OR", "WA")) +
  geom_boxplot() +
  labs(x="State", y="Yield") +
  ggtitle("Yield by State")
```

Research Question: How does each year's yield differ?

```{r}
strawberry.sur.wider %>%
  filter(Year %in% 1998:2011) %>%
  ggplot(aes(x=as.character(Year), y=`YIELD, MEASURED IN CWT / ACRE`, fill=as.character(Year))) +
  geom_boxplot() +
  labs(x="Year", y="Yield") +
  ggtitle("Yield by Year")

strawberry.sur.wider %>%
  filter(Year %in% 2012:2024) %>%
  ggplot(aes(x=as.character(Year), y=`YIELD, MEASURED IN CWT / ACRE`, fill=as.character(Year))) +
  geom_boxplot() +
  labs(x="Year", y="Yield") +
  ggtitle("Yield by Year") 
```

Research Question: How does the yield of each state change by year?

```{r}
strawberry.sur.wider %>%
  ggplot(aes(x=Year, y=`YIELD, MEASURED IN CWT / ACRE`, color=State)) +
  geom_line() +
  geom_point()+
  labs(x="Year", y="Yield") +
  ggtitle("Yield of Each State by Year")
```

```{r}
strawberry %>%
  distinct(Data.Item)
```

(This doesn't work, delete later )

```{r}
# Check for columns that contain very little information
# Amount of info is indicated by the number of non NA rows of each column
cols <- ncol(strawberry.sur.wider)
rows <- nrow(strawberry.sur.wider)
for (i in 1:cols){
  name <- names(strawberry.sur.wider)[i]
  notna <- sum(!is.na(strawberry.sur.wider[,i]))
  print(paste("Col No.:", i))
  print(name)
  print(notna/rows)
}
# Remove all columns whose non na rows are less than 50%
# They are columns whose names contain "UTILIZED", "SOLD", "PROCESSING"
strawberry.sur.wider <- strawberry.sur.wider %>%
  select(!contains("UTILIZED")) %>%
  select(!contains("SOLD")) %>%
  select(!contains("PROCESSING")) 

# Check if 
```
