---
title: "Assignment F Group 9 Report"
format: html
editor: visual
---

## Introduction

In this project, we will explore the data of strawberry agriculture in the United States obtained from USDA. Specifically, our research will focus on the strawberry yield, the ratio of crop grown to the unit area of land, of various states from 1998 to 2024.

Firstly, we will conduct some basic exploratory data analysis (EDA) on the survey data of strawberry yield of some selected states, trying to answer the following research questions:

1\) How did the yield of each state differ?

2\) How did the yield of each year differ?

3\) How did the yield of each state change across the years?

4\) Is there a correlation between yield and

Next, we will explore the census data on the strawberry yield of the entire U.S.,

Additionally, we will see what other insights about strawberry agriculture in the U.S. may be extracted from the data available to us.

## Set Up

```{r}
#| label: load libraries and set options
#| warning: false
#| message: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
library(ggplot2)
library(cowplot)
```

```{r}
# Define function: drop_one_value_col, which drops all columns that contain only one value
drop_one_value_col <- function(df){
  drop <- c()
  for (i in 1:ncol(df)){
    uni <- unique(df[,i])
    if (is.data.frame(uni)){
      unique_val <- nrow(unique(df[,i]))
    } else {
      unique_val <- length(unique(df[,i]))
    }

    
    
    if (unique_val == 1){
      drop <- c(drop, i)
    }
  }
  
  if (length(drop) != 0){
    df <- df[,-drop]
  }
  df
}
```

## EDA: Strawberry Yield of the Selected States

### Load and Clean the Data

```{r}
# Load the strawberrt data from csv file 
strawberry <- read.csv("strawberry_10oct25.csv") %>% 
  # Drop all single value column
  drop_one_value_col()

# Divide the data into two parts: survey & census
strawberry.sur <- strawberry %>%
  filter(Program=="SURVEY") 
strawberry.cen <- strawberry %>%
  filter(Program=="CENSUS")

```

Select data that is relevant to our research theme: yield

Since state census data contains nothing relevant to yield, so we will focus solely on the survey data for this section

```{r}
# Select survey data 
strawberry.sur <- strawberry %>%
  filter(Program=="SURVEY") %>%
  # Select Data.Item rows that contains information relevant to yield
  filter(grepl("ACRES|PRODUCTION|YIELD", Data.Item)) %>%
  # Drop all one value columns
  drop_one_value_col()

# Clean the data 
# Remove all "STRAWBERRIES - " from the column Data.Item, since they serve no purpose 
strawberry.sur$Data.Item <- str_remove_all(strawberry.sur$Data.Item, "STRAWBERRIES - ") 
# Remove all "STRAWBERRIES, " from the column Data.Item, since they serve no purpose 
strawberry.sur$Data.Item <- str_remove_all(strawberry.sur$Data.Item, "STRAWBERRIES, ") 

# Replace "(D)" with NA
strawberry.sur$Value[strawberry.sur$Value ==" (D)"] <- NA

# Turn all values into numeric
strawberry.sur$Value <- str_remove_all(strawberry.sur$Value, ",")
strawberry.sur$Value <- as.numeric(strawberry.sur$Value)
```

Reshape the data

```{r}
strawberry.sur.wider <- strawberry.sur %>% 
  pivot_wider(
    names_from = Data.Item,
    values_from = Value
    ) %>%
  # We do not consider "FORECAST" data, filter those rows out
  filter(!grepl("FORECAST", Period)) %>%
  # Arrange rows by state
  arrange(State)  %>%
  # Drop columns that contained "TON"
  select(!contains("TON")) %>%
  # Drop all one value columns
  drop_one_value_col()
```

### Question 1: How does each state's yield differ?

Here the bar chart shows the median strawberry yield of each state. We can see a clear difference between different states: the California has a median yield much higher than all other states, while New York has a median yield much lower than all other states.

```{r}
strawberry.sur.wider %>%
  group_by(State) %>%
  summarise(Median_Yield = median(`YIELD, MEASURED IN CWT / ACRE`, na.rm = TRUE)) %>%
  ggplot(aes(x = State, y = Median_Yield)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() + 
    labs(title = "Median Strawberry Yield by State",
         x = "State",
         y = "Median Yield (CWT / ACRE)") +
    theme_minimal()
```

Next, we will use boxplot to show the difference in strawberry yield between each state

```{r}
strawberry.sur.wider %>%
  ggplot(aes(x=State, y=`YIELD, MEASURED IN CWT / ACRE`, fill=State)) +
  scale_x_discrete(labels=c("CA", "FL", "NY", "NC", "OR", "WA")) +
  geom_boxplot() +
  labs(x="State", y="Yield (CWT/Acres)") +
  ggtitle("Yield by State")
```

It seems there is an clear differences in yield of strawberry between states. California overall has a yield much higher than any other state, while the yield of Florida, despite having a much lower comparing to California, is significantly higher than the rest of the states.

```{r}
strawberry.sur.wider %>%
  ggplot(aes(x=State, y=`ACRES PLANTED`, fill=State)) +
  scale_x_discrete(labels=c("CA", "FL", "NY", "NC", "OR", "WA")) +
  geom_boxplot() +
  labs(x="State", y="Yield (CWT/Acres)") +
  ggtitle("Yield by State")
```

```{r}
strawberry.sur.wider %>%
  ggplot(aes(x=State, y=`PRODUCTION, MEASURED IN CWT`, fill=State)) +
  scale_x_discrete(labels=c("CA", "FL", "NY", "NC", "OR", "WA")) +
  geom_boxplot() +
  labs(x="State", y="Yield (CWT/Acres)") +
  ggtitle("Yield by State")
```

### Question 2: How does each year's yield differ?

```{r}
strawberry.sur.wider %>%
  group_by(Year) %>%
  summarise(Median_Yield = median(`YIELD, MEASURED IN CWT / ACRE`, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = Median_Yield)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() + 
    labs(title = "Median Strawberry Yield by Year",
         x = "Year",
         y = "Median Yield (CWT / ACRE)") +
    theme_minimal()
```

```{r}
strawberry.sur.wider %>%
  filter(Year %in% 1998:2011) %>%
  ggplot(aes(x=as.character(Year), y=`YIELD, MEASURED IN CWT / ACRE`, fill=as.character(Year))) +
  geom_boxplot() +
  labs(x="Year", y="Yield (CWT/Acres)") +
  ggtitle("Yield by Year")

strawberry.sur.wider %>%
  filter(Year %in% 2012:2024) %>%
  ggplot(aes(x=as.character(Year), y=`YIELD, MEASURED IN CWT / ACRE`, fill=as.character(Year))) +
  geom_boxplot() +
  labs(x="Year", y="Yield (CWT/Acres)") +
  ggtitle("Yield by Year") 
```

From the graph, we can see that there is a great variation in the yield of each year. Also, there is a big jump in yield from 2018 to 2019. Judging by the previous graph, we conclude that the high variation is due to the fact that two of the states, California and Florida, have much higher yield comparing to the rest. For this reason, we decide to redo the plot while excluding the data from California and Florida.

```{r}
strawberry.sur.wider %>%
  group_by(Year) %>%
  filter(!State %in% c("CALIFORNIA", "FLORIDA")) %>%
  summarise(Median_Yield = median(`YIELD, MEASURED IN CWT / ACRE`, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = Median_Yield)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() + 
    labs(title = "Median Strawberry Yield by Year",
         subtitle = "Exclude CA, FL",
         x = "Year",
         y = "Median Yield (CWT / ACRE)") +
    theme_minimal()
```

```{r}
strawberry.sur.wider %>%
  filter(Year %in% 1998:2008, !State %in% c("CALIFORNIA", "FLORIDA")) %>%
  ggplot(aes(x=as.character(Year), y=`YIELD, MEASURED IN CWT / ACRE`, fill=as.character(Year))) +
  geom_boxplot() +
  labs(x="Year", y="Yield (CWT/Acres)") +
  ggtitle("Yield by Year (Exclude CA, FL)") +
  ylim(20,150)

strawberry.sur.wider %>%
  filter(Year %in% 2009:2024, !State %in% c("CALIFORNIA", "FLORIDA")) %>%
  ggplot(aes(x=as.character(Year), y=`YIELD, MEASURED IN CWT / ACRE`, fill=as.character(Year))) +
  geom_boxplot() + 
  labs(x="Year", y="Yield (CWT/Acres)") +
  ggtitle("Yield by Year (Exclude CA, FL)") +
  ylim(20,150)


```

Here we can see that the variation of yield between different years is much smaller. The new graph only includes year 1998 to 2018, because there is no data after 2018 for states other than California and Florida, this explains the "jump" from 2018 to 2019 in the previous graph: because since 2019, we only have data from California and Florida and these two states tend to have a much higher yield. Lastly, the yield seems to be quite stable across the years.

We will not conduct a separate analysis on the yearly yield of states California and Florida in this section. Since it does not make sense to use bar plot and boxplot when there are only two data point (one for California, one for Florida) for each year. For instead, the behavior of the yield of all states over time will be explored in the next section.

### Question 3: How does the yield of each state change by year?

```{r}
strawberry.sur.wider %>%
  ggplot(aes(x=Year, y=`YIELD, MEASURED IN CWT / ACRE`, color=State)) +
  geom_line() +
  geom_point()+
  labs(x="Year", y="Yield (CWT/Acres)") +
  ggtitle("Yield of Each State by Year")
```

From the first graph we can see that the strawberry yield of California has much higher variation. Despite this, the yield stays within the range between 500 and 800 CTW/Acres, which is still much higher than all other states. Florida also seems to have a fairly high yield (and variation). Althought it is significantly lower than California, it is still higher than all other states.The remaining states have a very similar yield which do not vary much.

From the second figure, we can see the overall trend of each state's strawberry yield separately. California has an increasing trend, despite the great variation. Florida on the other hand has a decreasing trend. Interestingly, we can see that as the yield decrease, the variation also decrease. The strawberry yield of all other states maintain a very stable trend, with low value and variation. We may hypothesize that states which are low in yield value also tend to have low variation.

### (Too complicated)

```{r}
#| warning: false
#| message: false

states <- unique(strawberry.sur.wider$State)
qqplots <- c()
plots <- c()
for (state in states){
  state.df <- strawberry.sur.wider %>% 
    filter(State==state) 
  plot <- state.df %>%
  ggplot(aes(x = `ACRES PLANTED`, y = `YIELD, MEASURED IN CWT / ACRE`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = state,
       x = "Acres Planted",
       y = "Yield (CWT/ACRE)") 
  plots <- c(plots, plot)
  
  qqplot <- state.df %>%
    ggplot(aes(x = `ACRES PLANTED`, y = `YIELD, MEASURED IN CWT / ACRE`)) +
    stat_qq() +
    stat_qq_line()
  qqplots <- c(qqplot, qqplots)
}

plot_grid(plotlist = plots, ncol=2)
plot_grid(plotlist = qqplots, ncol=2)
```

```{r}
#| warning: false
#| message: false

states <- unique(strawberry.sur.wider$State)
plots <- c()
for (state in states){
  state.df <- strawberry.sur.wider %>% 
    filter(State==state) %>%
    mutate(`PRODUCTION, MEASURED IN CWT` = `PRODUCTION, MEASURED IN CWT`/100)
  plot <- state.df %>%
  ggplot(aes(x = `PRODUCTION, MEASURED IN CWT`, y = `YIELD, MEASURED IN CWT / ACRE`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = state,
       x = "Production (100 CWT)",
       y = "Yield (CWT/ACRE)") 
  plots <- c(plots, plot)
}

plot_grid(plotlist = plots, ncol=2)
```

## EDA: Organic Strawberry Yield of the Entire U.S.

### Load and clean the data

```{r}

# Load the national census strawberry data from csv file 
strawberry.national <- read.csv("StrawberryCensusNational.csv") %>% 
  filter(!grepl("AREA GROWN", Domain)) %>% 
  filter(!grepl("OPERATIONS", Data.Item)) %>%
  filter(!grepl("SALES", Data.Item)) %>%
  # Drop all single value column
  drop_one_value_col()

# Remove redundant strings from Data.Item column
strawberry.national$Data.Item <- str_remove_all(strawberry.national$Data.Item, "STRAWBERRIES - ") 
strawberry.national$Data.Item <- str_remove_all(strawberry.national$Data.Item, "STRAWBERRIES, ") 


```

### EDC: Organic strawberry yield

```{r}
# Select and reorgnize all data relevant to the organic strawberry
strawberry.national.organic <- strawberry.national %>% 
  filter(grepl("ORGANIC", Domain)) %>%
  filter(Year>=2015 | grepl("(NOP USDA CERTIFIED & EXEMPT)", Domain.Category)) %>%
  select(Year, Data.Item, Domain.Category, Value) 
strawberry.national.organic$Data.Item <- str_remove_all(strawberry.national.organic$Data.Item, "ORGANIC - ") 

strawberry.national.organic <- strawberry.national.organic %>%
  pivot_wider(
    names_from = Data.Item,
    values_from = Value
    ) 

# Turn the acres value into numeric
strawberry.national.organic$`ACRES HARVESTED` <- str_remove_all(strawberry.national.organic$`ACRES HARVESTED`, ",")
strawberry.national.organic$`ACRES HARVESTED` <- as.numeric(strawberry.national.organic$`ACRES HARVESTED`)
# Turn the production value into numeric
strawberry.national.organic$`PRODUCTION, MEASURED IN CWT` <- str_remove_all(strawberry.national.organic$`PRODUCTION, MEASURED IN CWT`, ",")
strawberry.national.organic$`PRODUCTION, MEASURED IN CWT` <- as.numeric(strawberry.national.organic$`PRODUCTION, MEASURED IN CWT`)

# Select columns relevant to our analysis
strawberry.national.organic <- strawberry.national.organic %>% 
  mutate(`YIELD (CWT / ACRES)` = `PRODUCTION, MEASURED IN CWT`/`ACRES HARVESTED`) %>% select(-Domain.Category)

```

```{r}
strawberry.national.organic %>%
  ggplot(aes(x=Year, y=`YIELD (CWT / ACRES)`)) +
  geom_point() +
  labs(title="Organic Yield")
```

As shown in the graph above, we only have 5 data points in total. We conclude that there is not enough data to conduct meaningful analysis.

## Conclusion

### EDA: Strawberry yield of selected states from 1998 to 2024

We have some interesting findings from the exploratory data analysis of strawberry yield data of the selected states from 1998 to 2024.

1\) California has a significantly higher yield comparing to other states and is overall increasing.

2\) Florida has a yield much lower than California, but also much higher than the rest of the states. Though it seems to have a decreasing trend. Interestingly, as the yield of Florida decrease in value, its variation also decrease.

3\) All other states have similar yield with low variation and a stable trend.

Our research topics focus on columns containing the data of "acres", "total productions", "yield". There are other columns in the dataframe that are potentially useful for other research topics (ex., comparing the productions between "Fresh Market", "Processing", "Not Sold", "Utilized"). But, due to the vast amount of missing values in those columns, we conclude that there is not much insight we can extract from them.

Finally, our analysis lead to some new research questions for future study. Here are some examples:

1\) What factors make the strawberry yield of California and Florida so high? What distinguish them from the other states selected for this study?

2\) Does strawberry yield value have a negative correlation to low variation?

3\) How does the strawberry yield relate to other matrices that determine the success of strawberry agriculture in a state?

### EDA: Organic vs Conventional strawberry yield of the entire U.S.

Because we do not have sufficient data , we do not have any meaningful findings about the difference between organic strawberry and conventional strawberry in terms of yield. On the good side, it provides us a good opportunity to practice data cleaning.
